Bounds, Semantics/Pragmatics-Divide, and Evolution

### 2016-08-02 paper/paper.tex
paper section 2 draft update

\begin{document}
\maketitle
\tableofcontents
\section{The semantics-pragmatics divide}\label{sec:introduction}

@@

Prima facie, it is puzzling that {\em Bill read five books} does not semantically rule out its more informative or ``stronger'' alternatives. More poignantly, would it not serve language users better if weak(er) expressions such as {\em warm}, {\em or}, {\em some} or {\em big} were truth-conditionally incompatible with stronger alternatives such as, respectively, {\em hot}, {\em and}, {\em all} and {\em huge}? After all, analogously to the example sketched above, this is what a pragmatically enriched interpretation yields; an upper-bound that rules out these alternatives through a so-called scalar inference \citep{horn:1972,gazdar:1979}. Once a distinction between semantics and pragmatics is drawn, similar questions can be posed for other types of pragmatic inferences, such as ignorance inferences as and manner inferences.

@@

\section{Conveying upper-bounds}
\hl{start bottom up -- from the behavior of agents to the overall dynamics. Here, begin with language use and introduce scalar inferences on the way as the object of study we will focus on here}

@@

\section{Upper-bounds in weak scalar alternatives}
\hl{phenomenon description}

\subsection{Analysis}

>>>>

\begin{document}
\maketitle
\section{The semantics-pragmatics divide}\label{sec:introduction}

@@

Prima facie, it is puzzling that {\em Bill read five books} does not semantically rule out its more informative or ``stronger'' alternatives. More poignantly, would it not serve language users better if weak(er) expressions such as {\em warm}, {\em or}, {\em some} or {\em big} were truth-conditionally incompatible with stronger alternatives such as, respectively, {\em hot}, {\em and}, {\em all} and {\em huge}? After all, analogously to the example sketched above, this is what a pragmatically enriched interpretation yields; an upper-bound that rules out these alternatives through a so-called scalar inference \citep{horn:1972,gazdar:1979}. The present investigation focuses on the lack of lexicalization of such upper-bounds in natural languages. That is, we seek to provide an explanation for the selection of the particular semantics of scalar expressions, which in turn hinges on the semantics-pragmatics distinction for what is ultimately conveyed to interlocutors. However, once a distinction between semantics and pragmatics is drawn, similar questions can be posed for other types of pragmatic inferences, such as ignorance and manner inferences.

@@

\section{Conveying upper-bounds}
Scalar inferences involve the pragmatic derivation of an upper-bound for weak scalar expressions provided the hearer assumes the speaker to be knowledgeable and cooperative. A scale is an informativity induced ordering of relevant alternatives. Here, informativity is understood as entailment. For instance, {\em all students came} entails {\em most students came}, which in turn entails {\em some students came}. In this sense, {\em some} is weaker than {\em all}. The inference derived from utterances involving weaker alternatives is that stronger ones do not hold, e.g. {\em some students came} may be taken to convey that {\em not all students came}.

\hl{say something about this pattern being widespread across languages. Give examples of classes of expressions in English, as alluded above (numerals, scalar adjectives, quantifiers)}

Framed in Gricean terms \citep{grice:1975}, the hearer reasons about the speaker's choice of a weak alternative over a stronger one; had the speaker known that a stronger alternatives holds, she would have said so as this would have been more informative, since she did not, the hearer can infer that the stronger alternatives do not hold. Analogously, a speaker who reasons about her addressee may be rely on her to derive this inference. In this way, the strengthened, upper-bounded, state of affairs can be conveyed without codifiying the bound explicitly in the semantics.

\paragraph{Function-based explanations.} Two ingredients are key to this derivation; (the assumptions of) cooperation and knowledge about the issue at hand. It is not difficult to imagine cases were one or both are not given. The speaker may (be assumed to) want to withhold information about the students' attendance, or may have left early without being able to verify the attendance to a satisfactory degree.

\paragraph{Learning-based explanations.}

\hl{start bottom up -- from the behavior of agents to the overall dynamics. Here, begin with language use and introduce scalar inferences on the way as the object of study we will focus on here}

@@

\section{Analysis}
